---
title: "Diffusion"
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(deSolve)
library(here)
```

# R implementation

```{r}
source(here("R/diffusion.R"))

# run our diffusion model (iterative difference equation) with initial concentration of 10, for 8 timestep (size 1m), and 10 space steps (size 1s)
# using diffusion parameters 0.5 s/m2, 10 m2
result <- diff1(initialC = 10, nx = 10, dx = 1, nt = 8, dt = 1, D = 0.5, area = 10)

# a list is returned with our 3 data frames for concentration (conc), qin and qout
result

# used filled contour to plot results
head(result$conc)
filled.contour(result$conc, xlab = "Time", ylab = "Distance")

# or if you prefer this orientation (Distance on x axis)
filled.contour(t(result$conc), ylab = "Time", xlab = "Distance")
```

# Change parameters (diffusivity D, and space and time steps (dx, dt))

```{r}
# changes diffusivity and other parameters particularly
# diffusivity, dx and dt
res <- diff1(initialC = 10, nx = 10, dx = 1, nt = 8, dt = 1, D = 0.5, area = 10)

filled.contour(res$conc, xlab = "Time", ylab = "Distance")


# we can also see how much material moved from place to place each time step
filled.contour(res$qin, xlab = "Time", ylab = "Distance")


# play with time step, space step and parameters
```

pollutant moving through time and space

when diffusion gets real high, or other combos of parameters, we get these crazy looking triangles that do not make much physical sense - something is wrong. unstable equilibria?? oscilating perhaps??
in this case, either time step is too small with too big space step, or vice versa 
its moving too much from one space to another, changing the relationship between cells

time steps are too coarse based on diffusivity



# 1. Play with time step, space step and parameters
# 2. Design different ways to summarize and visualize output 

```{r}
source(here("R/diff_sum.R"))

diff_summ(res)

```





